---
title: "Series 2"
author: "Christian Holberg"
date: "9/23/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Exercise 1

Loading data

```{r}
traffic_data <- tibble(
  year = seq(2006, 2016, 1),
  traffic_jams = c(112, 109, 91, 98, 139, 150, 168, 149, 149, 135, 140)
)
head(traffic_data)
```

Scatter plot of data

```{r}
ggplot(traffic_data, aes(x = year, y = traffic_jams)) + geom_point()
```


### a)

Defining function for computing the Running Mean Smoother

```{r}
myRMS <- function(start, finish){
  window <- 3
  step <- 1
  num_steps <- floor((finish-start)/step)
  running_mean <- numeric(num_steps+1)
  for (i in 1:(num_steps+1)){
    center <- start + i*step - 1
    y_range <- traffic_data$traffic_jams[(traffic_data$year <= center + window/2 &
                                  traffic_data$year >= center - window/2)]
    running_mean[i] <- mean(y_range)
  }
  return(running_mean)
}
```

Using Running Mean Smoother and drawing solution

```{r}
ggplot(traffic_data, aes(x = year)) + 
  geom_point(aes(y = traffic_jams)) +
  geom_line(aes(y = myRMS(2006, 2016)))
```

### b)

Computing the running mean with the build in `ksmooth()` function and plotting it in the scatter plot.

```{r}
running_mean <- with(traffic_data, ksmooth(year, traffic_jams, "box", 3.0, x.points = 2006:2016))
ggplot(traffic_data, aes(x = year)) + 
  geom_point(aes(y = traffic_jams)) +
  geom_line(aes(y = running_mean$y))
```

They look exactly the same, which was to be expected.

### c)

If the (unnormalized) density of a Gaussian distribution with standard deviation \(\sigma=2\) and mean \(\mu\) is given by

$$exp\left(-\frac{(\mu-x)^2}{2\sigma^2}\right)=exp\left(-\frac{(x-x_i)^2}{8}\right)$$

then the weights of the corresponding Gaussian Kernel Smoother must be

$$w_i = exp\left(-\frac{(x-x_i)^2}{\lambda}\right)$$
with \(\lambda = 8\). Below is a function for computing the Gaussian Kernel Smoother with \(\lambda = 8\)

```{r}
myGKS <- function(start, finish, num_steps){
  step <- (finish-start)/num_steps
  running_mean <- numeric(num_steps+1)
  for (i in 1:(num_steps+1)){
    center <- start + (i-1)*step
    w <- exp(-((center - traffic_data$year)^2)/8)
    running_mean[i] <- (w %*% traffic_data$traffic_jams)/sum(w)
  }
  return(running_mean)
}
```

Computing and drawing solution in the scatter plot.

```{r}
ggplot() + 
  geom_point(aes(x = year, y = traffic_jams), data = traffic_data) +
  geom_line(aes(x = seq(2006, 2016, 0.1), y = myGKS(2006, 2016, 100)))
```

### d)

Now using the built in `ksmooth()` function. From the `ksmooth()` documentation it follows that a bandwidth of `qnorm(0.75)*2*4` should correspond to the Gaussian with a standard deviation of 2.

```{r}
running_mean <- with(traffic_data, ksmooth(year, traffic_jams, "normal", bandwidth = qnorm(0.75)*2*4, 
                                           x.points = seq(2006, 2016, 0.1)))
ggplot() + 
  geom_point(aes(x = year, y = traffic_jams), data = traffic_data) +
  geom_line(aes(x = running_mean$x, y = running_mean$y))
```

### e)

LOESS smoothers for different values of `degree` and `span`.

```{r}
loess1 <- with(traffic_data, loess.smooth(year, traffic_jams, 2/3, 1, "gaussian", evaluation = 101))
loess2 <- with(traffic_data, loess.smooth(year, traffic_jams, 1, 1, "gaussian", evaluation = 101))
loess3 <- with(traffic_data, loess.smooth(year, traffic_jams, 2/3, 2, "gaussian", evaluation = 101))

ggplot() +
  geom_point(aes(x = year, y = traffic_jams), data = traffic_data) +
  geom_line(aes(x = loess1$x, y = loess1$y), col = "red") +
  geom_line(aes(x = loess2$x, y = loess2$y), col = "blue") +
  geom_line(aes(x = loess3$x, y = loess3$y), col = "green")
```

All three smoothers in one plot